---
title: "Newdata"
author: "Chang Lu"
date: "2025-03-27"
output: pdf_document
---


```{r}
library(readr)
library(dplyr)
library(lubridate)
library(stringr)
library(purrr)
library(readxl)
```

```{r, include=FALSE}
mams_df   <- read_excel("Click Times for MAMS and Dental Students.xlsx", sheet = "MAMS Students")
dental_df <- read_excel("Click Times for MAMS and Dental Students.xlsx", sheet = "Dental Students")

mams_df   <- mams_df[-1, ]
dental_df <- dental_df[-1, ]

# nrow(mams_df)   
# nrow(dental_df)   

dental_df <- dental_df %>% filter(Progress != "Progress")

#nrow(dental_df)


combined_df <- bind_rows(mams_df, dental_df)
summary(combined_df)
```

```{r}
# Computing Total_Quit_Time_Computed
submit_time_cols <- names(combined_df)[str_detect(names(combined_df), "-T_Page Submit")]

combined_df$Total_Quit_Time_Computed <- combined_df %>%
  select(all_of(submit_time_cols)) %>%
  mutate_all(as.numeric) %>%
  rowSums(na.rm = TRUE)


combined_df %>% select(...1, Total_Quit_Time_Computed) %>% head()
length(combined_df$Total_Quit_Time_Computed)
```


```{r}
# Rename demographic groups
combined_df <- combined_df %>%
  rename(
    Accommodations       = Q21,
    Gender               = Q22,
    Ethnicity            = Q23,
    English_Proficiency  = Q24,
    Country_You          = Q25_1,
    Country_Mother       = Q25_2,
    Country_Father       = Q25_3,
    Home_Language        = Q69
  ) %>%
  select(-Q68)  # Drop Q68 for a few people answered


```


### Normality Check


#### Accomodations

```{r}
# # Accomodations
# group_yes <- combined_df %>% filter(Accommodations == "Yes") %>% pull(Total_Quit_Time_Computed)
# group_no  <- combined_df %>% filter(Accommodations == "No") %>% pull(Total_Quit_Time_Computed)
# 
# # Shapiro-Wilk test
# shapiro.test(group_yes)
# shapiro.test(group_no)
```
```{r}
# par(mfrow = c(2, 2))  # Plot layout
# 
# # Histogram
# hist(group_yes, main = "Accommodations: YES", xlab = "Total Time", col = "skyblue")
# hist(group_no, main = "Accommodations: NO", xlab = "Total Time", col = "salmon")
# 
# # Q-Q plots
# qqnorm(group_yes); qqline(group_yes, col = "blue")
# qqnorm(group_no); qqline(group_no, col = "red")
```

Error in shapiro.test(group_yes) : sample size must be between 3 and 5000 -> Mann–Whitney U test

#### Gender
```{r}
# Group sizes
table(combined_df$Gender)

# Normality check
by(combined_df$Total_Quit_Time_Computed, combined_df$Gender, shapiro.test)
```

We tested whether total quit time differed by gender. The Female group violated the assumption of normality (Shapiro-Wilk p < 0.001), so we used a non-parametric Mann–Whitney U test.

#### Race

```{r}
table(combined_df$Ethnicity)

# Kruskal-Wallis test (non-parametric ANOVA)
kruskal.test(Total_Quit_Time_Computed ~ Ethnicity, data = combined_df)
```

A Kruskal-Wallis test revealed no significant difference in total quit time across ethnic groups ($\chi^2(4)$ = 1.06, p = 0.901).

#### English Proficiency

```{r}
# Clean and recode English proficiency
combined_df <- combined_df %>%
  mutate(English_Proficiency = case_when(
    str_trim(English_Proficiency) %in% c("Native", "Native Speaker") ~ "Native",
    str_trim(English_Proficiency) == "Fluent" ~ "Fluent",
    TRUE ~ English_Proficiency  # keep as-is in case there's other unexpected values
  ))

table(combined_df$English_Proficiency)

wilcox.test(Total_Quit_Time_Computed ~ English_Proficiency, data = combined_df)
```

A Wilcoxon rank-sum test found no significant difference in total quit time between Fluent and Native English speakers (W = 113, p = 0.89)

#### Born in US

```{r}
# You
combined_df <- combined_df %>%
  mutate(Born_in_US = ifelse(Country_You == "United States", "Yes", "No"))

wilcox.test(Total_Quit_Time_Computed ~ Born_in_US, data = combined_df)
```

A Wilcoxon rank-sum test found no significant difference in total quit time between students born in the U.S. and those born elsewhere (W = 92, p = 0.73).


```{r}
# Mother
combined_df <- combined_df %>%
  mutate(Born_in_US_M = ifelse(Country_Mother == "United States", "Yes", "No"))

wilcox.test(Total_Quit_Time_Computed ~ Born_in_US_M, data = combined_df)
```

A Wilcoxon rank-sum test comparing total quit time by mother's country of birth showed a trend toward significance (W = 156, p = 0.093), but did not reach conventional significance levels. It may need larger sample size to support.


```{r}
# Father
combined_df <- combined_df %>%
  mutate(Born_in_US_F = ifelse(Country_Father == "United States", "Yes", "No"))

wilcox.test(Total_Quit_Time_Computed ~ Born_in_US_F, data = combined_df)
```

A Wilcoxon rank-sum test showed no significant difference in total quit time based on father’s country of birth (W = 149, p = 0.211).

#### Home Language

```{r}
combined_df <- combined_df %>%
  mutate(English_Home = ifelse(Home_Language == "English", "Yes", "No"))

wilcox.test(Total_Quit_Time_Computed ~ English_Home, data = combined_df)
```

A Wilcoxon rank-sum test showed no significant difference in total quit time between students who spoke English at home and those who did not (W = 152, p = 0.130).

### To qiuyi

Hi Qiuyi, when comparing total quit time across demographic groups, here's how to decide between using a t-test or the Mann–Whitney U test:


Use t-test if:
Both groups have at least 10 observations and the distribution of total quit time is approximately normal in both groups (check via Shapiro-Wilk or Q-Q plot)(I didnt draw plots) \\

Use Mann–Whitney U test if:
1.One or both groups violate normality

2.Sample size is small (especially < 10)


