---
title: "Newdata"
author: "Chang Lu"
date: "2025-03-27"
output: pdf_document
---


```{r}
library(readr)
library(dplyr)
library(lubridate)
library(stringr)
library(purrr)
library(readxl)
```

```{r, include=FALSE}
mams_df   <- read_excel("Click Times for MAMS and Dental Students.xlsx", sheet = "MAMS Students")
dental_df <- read_excel("Click Times for MAMS and Dental Students.xlsx", sheet = "Dental Students")

mams_df   <- mams_df[-1, ]
dental_df <- dental_df[-1, ]

# nrow(mams_df)   
# nrow(dental_df)   

dental_df <- dental_df %>% filter(Progress != "Progress")

#nrow(dental_df)


combined_df <- bind_rows(mams_df, dental_df)
summary(combined_df)
```

```{r}
# Computing Total_Quit_Time_Computed
submit_time_cols <- names(combined_df)[str_detect(names(combined_df), "-T_Page Submit")]

combined_df$Total_Quit_Time_Computed <- combined_df %>%
  select(all_of(submit_time_cols)) %>%
  mutate_all(as.numeric) %>%
  rowSums(na.rm = TRUE)


combined_df %>% select(...1, Total_Quit_Time_Computed) %>% head()
length(combined_df$Total_Quit_Time_Computed)
```


```{r}
# Rename demographic groups
combined_df <- combined_df %>%
  rename(
    Accommodations       = Q21,
    Gender               = Q22,
    Ethnicity            = Q23,
    English_Proficiency  = Q24,
    Country_You          = Q25_1,
    Country_Mother       = Q25_2,
    Country_Father       = Q25_3,
    Home_Language        = Q69
  ) %>%
  select(-Q68)  # Drop Q68 for a few people answered


```
```{r}
combined_df
```


### Normality Check


#### Accomodations

```{r}
# # Accomodations
# group_yes <- combined_df %>% filter(Accommodations == "Yes") %>% pull(Total_Quit_Time_Computed)
# group_no  <- combined_df %>% filter(Accommodations == "No") %>% pull(Total_Quit_Time_Computed)
# 
# # Shapiro-Wilk test
# shapiro.test(group_yes)
# shapiro.test(group_no)
```
```{r}
# par(mfrow = c(2, 2))  # Plot layout
# 
# # Histogram
# hist(group_yes, main = "Accommodations: YES", xlab = "Total Time", col = "skyblue")
# hist(group_no, main = "Accommodations: NO", xlab = "Total Time", col = "salmon")
# 
# # Q-Q plots
# qqnorm(group_yes); qqline(group_yes, col = "blue")
# qqnorm(group_no); qqline(group_no, col = "red")
```

Error in shapiro.test(group_yes) : sample size must be between 3 and 5000 -> Mann–Whitney U test

#### Gender
```{r}
# Group sizes
table(combined_df$Gender)

# Normality check
by(combined_df$Total_Quit_Time_Computed, combined_df$Gender, shapiro.test)
```

We tested whether total quit time differed by gender. The Female group violated the assumption of normality (Shapiro-Wilk p < 0.001), so we used a non-parametric Mann–Whitney U test.

#### Race

```{r}
table(combined_df$Ethnicity)

# Kruskal-Wallis test (non-parametric ANOVA)
kruskal.test(Total_Quit_Time_Computed ~ Ethnicity, data = combined_df)
```

A Kruskal-Wallis test revealed no significant difference in total quit time across ethnic groups ($\chi^2(4)$ = 1.06, p = 0.901).

#### English Proficiency

```{r}
# Clean and recode English proficiency
combined_df <- combined_df %>%
  mutate(English_Proficiency = case_when(
    str_trim(English_Proficiency) %in% c("Native", "Native Speaker") ~ "Native",
    str_trim(English_Proficiency) == "Fluent" ~ "Fluent",
    TRUE ~ English_Proficiency  # keep as-is in case there's other unexpected values
  ))

table(combined_df$English_Proficiency)

wilcox.test(Total_Quit_Time_Computed ~ English_Proficiency, data = combined_df)
```

A Wilcoxon rank-sum test found no significant difference in total quit time between Fluent and Native English speakers (W = 113, p = 0.89)

#### Born in US

```{r}
# You
combined_df <- combined_df %>%
  mutate(Born_in_US = ifelse(Country_You == "United States", "Yes", "No"))

wilcox.test(Total_Quit_Time_Computed ~ Born_in_US, data = combined_df)
```

A Wilcoxon rank-sum test found no significant difference in total quit time between students born in the U.S. and those born elsewhere (W = 92, p = 0.73).


```{r}
# Mother
combined_df <- combined_df %>%
  mutate(Born_in_US_M = ifelse(Country_Mother == "United States", "Yes", "No"))

wilcox.test(Total_Quit_Time_Computed ~ Born_in_US_M, data = combined_df)
```

A Wilcoxon rank-sum test comparing total quit time by mother's country of birth showed a trend toward significance (W = 156, p = 0.093), but did not reach conventional significance levels. It may need larger sample size to support.


```{r}
# Father
combined_df <- combined_df %>%
  mutate(Born_in_US_F = ifelse(Country_Father == "United States", "Yes", "No"))

wilcox.test(Total_Quit_Time_Computed ~ Born_in_US_F, data = combined_df)
```

A Wilcoxon rank-sum test showed no significant difference in total quit time based on father’s country of birth (W = 149, p = 0.211).

#### Home Language

```{r}
combined_df <- combined_df %>%
  mutate(English_Home = ifelse(Home_Language == "English", "Yes", "No"))

wilcox.test(Total_Quit_Time_Computed ~ English_Home, data = combined_df)
```

A Wilcoxon rank-sum test showed no significant difference in total quit time between students who spoke English at home and those who did not (W = 152, p = 0.130).

### To qiuyi

Hi Qiuyi, when comparing total quit time across demographic groups, here's how to decide between using a t-test or the Mann–Whitney U test:


Use t-test if:
Both groups have at least 10 observations and the distribution of total quit time is approximately normal in both groups (check via Shapiro-Wilk or Q-Q plot)(I didnt draw plots) \\

Use Mann–Whitney U test if:
1.One or both groups violate normality

2.Sample size is small (especially < 10)

### T-test and U-test

Due to the sample size, we decide to use both T-test and U-test to check each variable. The reason is that even the specific variable corresponding to the normality check but the small sample size leads to a large variance so we cannot believe the simple result.

## 1.English Proficiency

```{r}
group1 <- combined_df$`Total_Quit_Time_Computed`[combined_df$`English_Proficiency` == 'Fluent']
group2 <- combined_df$`Total_Quit_Time_Computed`[combined_df$`English_Proficiency` == 'Native']
t_test_result <- t.test(group1, group2, var.equal = TRUE)  
print(t_test_result)
```
```{r}
u_test_english <- wilcox.test(combined_df$Total_Quit_Time_Computed ~ combined_df$English_Proficiency, exact = FALSE)

print(u_test_english)
```

## 2.Race
```{r}
group1 <- combined_df$Total_Quit_Time_Computed[combined_df$Ethnicity == 'White']
group2 <- combined_df$Total_Quit_Time_Computed[combined_df$Ethnicity != 'White']

t_Race <- t.test(group1, group2, var.equal = TRUE)

print(t_Race)

```
```{r}
u_test_race <- wilcox.test(combined_df$Total_Quit_Time_Computed ~ (combined_df$Ethnicity == 'White'), exact = FALSE)

print(u_test_race)

```

## 3.Country_you
```{r}
group1 <- combined_df$Total_Quit_Time_Computed[combined_df$Country_You == 'United States']
group2 <- combined_df$Total_Quit_Time_Computed[combined_df$Country_You != 'United States']

t_Country <- t.test(group1, group2, var.equal = TRUE)

print(t_Country)
```
```{r}
u_test_country <- wilcox.test(group1, group2, exact = FALSE)

print(u_test_country)

```
## 4.Country Mother

```{r}
group1 <- combined_df$Total_Quit_Time_Computed[combined_df$Country_Mother == 'United States']
group2 <- combined_df$Total_Quit_Time_Computed[combined_df$Country_Mother != 'United States']

t_Mother <- t.test(group1, group2, var.equal = TRUE)

print(t_Mother)

```
```{r}
u_test_mother <- wilcox.test(group1, group2, exact = FALSE)

print(u_test_mother)
```
## 5.Country Father
```{r}
group1 <- combined_df$Total_Quit_Time_Computed[combined_df$Country_Father == 'United States']
group2 <- combined_df$Total_Quit_Time_Computed[combined_df$Country_Father != 'United States']

t_Father <- t.test(group1, group2, var.equal = TRUE)

print(t_Father)

```

```{r}
u_test_father <- wilcox.test(group1, group2, exact = FALSE)

print(u_test_father)

```


## 6.Home Language
```{r}
group1 <- combined_df$Total_Quit_Time_Computed[combined_df$Home_Language == 'English']
group2 <- combined_df$Total_Quit_Time_Computed[combined_df$Home_Language != 'English']

t_HomeLang <- t.test(group1, group2, var.equal = TRUE)

print(t_HomeLang)
```
```{r}
u_test_homeLang <- wilcox.test(group1, group2, exact = FALSE)

print(u_test_homeLang)

```
## Gender
```{r}
group1 <- combined_df$Total_Quit_Time_Computed[combined_df$Gender == 'Female']
group2 <- combined_df$Total_Quit_Time_Computed[combined_df$Gender != 'Female']

t_Gender <- t.test(group1, group2, var.equal = TRUE)

print(t_Gender)
```

```{r}
u_test_Gender <- wilcox.test(group1, group2, exact = FALSE)

print(u_test_Gender)

```
### Conclusion
So the final conclusion is:
```{r}
library(knitr)
data1 <- data.frame(
  Variable = c("English Proficiency", "Race", "Country You", 
               "Country Mother", "Country Father", "Home Language", "Gender"),
  `T-test P value` = c(0.4377, 0.9885, 0.6903, 0.6234, 0.7697, 1, 0.555),
  `U-test P value` = c(0.8886, 0.3267, 0.7231, 0.09237, 0.2073, 0.2073, 0.2497)
)

kable(data1, caption = "P-values from T-test and U-test for different variables")
```
The conclusion is that almost all variables they don't have such a huge significant difference in the aspect of quiz_time.
