---
title: "T_test"
author: "Qiuyi Feng"
date: "2025-02-20"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Since U-test and T-test only focus on binary data, so some category data like 'Age Arrive USA', continous data 'quiz time(categorized by 10 minutes)' don't work.

1. quiz time(categorized by 10 minutes)

2. English Proficiency

3. Race

4. Home Language

#### Mann-Whitney U Test

1. Accomodations

2. Age Arrive USA

3. Sex

4. Born in USA
```{r}
library(readxl)
df <- read_excel("MAMS and Dental Combined for Analysis.xlsx", sheet = "Sheet1")
df
summary(df)
```

```{r}
df_students <- df[2:32, ]

head(df_students)
# Define the new column names for the questions
question_names <- paste0("Q", 1:20)  # Generates Q1, Q2, ..., Q20

# Rename only the question columns
colnames(df_students)[2:21] <- question_names

head(df_students)
colnames(df_students)
```

```{r}

# Convert value fromat to numeric
df_students$`Total Score` <- as.numeric(df_students$`Total Score`)
df_students$`Quiz Time` <- as.numeric(df_students$`Quiz Time`)
df_students$Accomodations <- as.numeric(df_students$Accomodations)
df_students$Sex <- as.numeric(df_students$Sex)
df_students$`Race/Ethnicity` <- as.numeric(df_students$`Race/Ethnicity`)
df_students$`English Proficiency` <- as.numeric(df_students$`English Proficiency`)
df_students$`Born USA` <- as.numeric(df_students$`Born USA`)
df_students$`Home Language` <- as.numeric(df_students$`Home Language`)
df_students$`Age arrive USA`<- as.numeric(df_students$`Age arrive USA`)


# Convert Quiz Time from seconds to minutes
df_students$`Quiz Time Minutes` <- df_students$`Quiz Time` / 60 # for later test
# Create Quiz Time Bins
df_students$`Quiz Time Group` <- cut(df_students$`Quiz Time Minutes`, 
                                   breaks = c(0, 10, 20, Inf), 
                                   labels = c("0-10 min", "10-20 min", "20+ min"),
                                   include.lowest = TRUE)

# Check new groups
table(df_students$`Quiz Time Group`)


# Verify changes
str(df_students)
```


T-test
1.English Proiciency
```{r}
group1 <- df_students$`Total Score`[df_students$`English Proficiency` == 0]
group2 <- df_students$`Total Score`[df_students$`English Proficiency` == 1]

t_test_result <- t.test(group1, group2, var.equal = TRUE)  
print(t_test_result)
```

2.Race
```{r}
group1 <- df_students$`Total Score`[df_students$`Race/Ethnicity` == 0]
group2 <- df_students$`Total Score`[df_students$`Race/Ethnicity` == 1]

t_Race <- t.test(group1, group2, var.equal = TRUE)  
print(t_Race)
```
3.Home Language
```{r}
group1 <- df_students$`Total Score`[df_students$`Home Language` == 0]
group2 <- df_students$`Total Score`[df_students$`Home Language` == 1]

t_HomeLanguage <- t.test(group1, group2, var.equal = TRUE)  
print(t_HomeLanguage)
```

U-test
1.Sex
```{r}
u_test_sex <- wilcox.test(df_students$`Total Score` ~ df_students$Sex, exact = FALSE)

print(u_test_sex)

```
Sex is 0.139, which is not verified to the U test from client.

2.Born USA
```{r}

u_test_born <- wilcox.test(df_students$`Total Score` ~ df_students$`Born USA`, exact = FALSE)

print(u_test_born)

```
The p-value of Born in USA is also different 
3.Accomodations
```{r}
u_test_accomodations <- wilcox.test(df_students$`Total Score` ~ df_students$Accomodations, exact = FALSE)

print(u_test_accomodations)

```

We can find in all those different grouping method, the P-value is larger than 0.05, which means  there is no such a huge difference between diffient groups.
